Automated Metadata Generation for Fish Specimen Image Collections

### 2021-03-25 paper/paper.tex
Added more data and small tweaks

Placeholder\todo{A paragraph about training will go here}
\subsection{Pixel Analysis}
The masks and bounding boxes produced by \verb|detectron| are generally quite good, although they are almost never perfect. They almost always either include additional background as part of the fish, or clip part(s) of the fish outside the bounding box. To solve this shortcoming, we utilize pixel analysis methods commonly found in image informatics.

@@

To overcome this, the threshold value needs either scaled up or down depending on if the background is lighter or darker than the fish. For our current dataset, the background is always lighter (i.e.\ closer to 255), so the threshold value needs scaled up to include more of the image as foreground. Precisely how much to scale this value by is difficult to determine. For optimal results it is dependent on the level of contrast between the background and foreground, which can be affected by both how washed out the image is and how light or dark the pigmentation of the fish is. To take this into account, we measure the mean intensity of the foreground and background, then use these intensity values and the difference between them to scale the threshold value. Even with these values in hand it is not entirely clear how to scale the threshold, but we have found\ldots{}\todo{This is still somewhat indeterminate, need to pick a definitive technique, I've tried a few}

@@

\begin{table}[H]
    \centering
    \caption{Metadata properties}
    \label{tab:properties}
    \begin{tabular}{cccp{0.5\linewidth}}
        \toprule
        \textbf{Property} & \textbf{Association} & \textbf{Type} & \textbf{Explanation}\\
        \midrule
        \verb|has_fish| & Overall Image & Boolean & Whether a fish was found in the image.\\
        \verb|fish_count| & Overall Image & Integer & The quantity of fish present.\\
        \verb|has_ruler| & Overall Image & Boolean & Whether a ruler was found in the image.\\
        \verb|ruler_bbox| & Overall Image & 4 Tuple & The bounding box of the ruler (if found).\\
        \verb|scale| & Overall Image & Float & The scale of the image in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|bbox| & Per Fish & 4 Tuple & The top left and bottom right coordinates of what define the bounding box for a fish.\\
        \verb|background.mean| & Per Fish & Float & The mean intensity of the background within a given fish's bounding box.\\
        \verb|background.std| & Per Fish & Float & The standard deviation of the background within a given fish's bounding box.\\
        \verb|foreground.mean| & Per Fish & Float & The mean intensity of the foreground within a given fish's bounding box.\\
        \verb|foreground.std| & Per Fish & Float & The standard deviation of the foreground within a given fish's bounding box.\\
        \verb|centroid| & Per Fish & 4 Tuple & The centroid of a given fish's bitmask.\\
        \verb|primary_axis| & Per Fish & 2D Vector & The unit length primary axis (eigenvector) for the bitmask of a given fish.\\
        \verb|clock_value| & Per Fish & Integer & A fish's primary axis converted into an integer ``clock value'' between 1 and 12.\\
        \verb|length| & Per Fish & Float & The length of a fish in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|mask| & Per Fish & 2D Matrix & The bitmask of a fish in 0's and 1's.\\
        \verb|score| & Per Fish & Float & The percent confidence score output by \verb|detectron| for a given fish.\\
        \verb|has_eye| & Per Fish & Boolean & Whether an eye was found for a given fish.\\
        \verb|eye_center| & Per Fish & 2 Tuple & The centroid of a fish's eye.\\
        \verb|side| & Per Fish & String & The side (i.e.\ \verb|'left'| or \verb|'right'|) of the fish that is facing the camera (dependent on finding its eye).\\
      \bottomrule
\end{tabular}
\end{table}

@@

My hand checking of the ''wrong wrong'' fish:

INHS_FISH_101109.jpg: 9, 11, I'm wrong although it's fairly curved
INHS_FISH_56818.jpg: 4, 2, I'm wrong although there's a tag that is likely messing things up
INHS_FISH_57607.jpg: 3, 9, I'm wrong
INHS_FISH_57001.jpg: 4, 2, I'm wrong although there's a tag that is likely messing things up
INHS_FISH_63953.jpg: 3, 9, I'm wrong
INHS_FISH_51100.jpg: 3, 9, I'm wrong
INHS_FISH_80528.jpg: 3, 8, I'm wrong
INHS_FISH_6370.jpg: 3, 9, I'm wrong
INHS_FISH_54157.jpg: 9, 7, I'm arguably wrong although it's very curved
INHS_FISH_80987.jpg: 9, 3, I'm wrong although this one is a fairly atypical image

>>>>

\begin{verbatim}
|  category  | #instances   |  category  | #instances   |  category  | #instances   |
|:----------:|:-------------|:----------:|:-------------|:----------:|:-------------|
|    fish    | 297          |   ruler    | 1496         |    eye     | 456          |
|    two     | 100          |   three    | 100          |            |              |
|   total    | 2449         |            |              |            |              |

Iterations: 100,000

Initial: {"data_time": 0.004866079999999329, "fast_rcnn/cls_accuracy": 0.115234375,
"fast_rcnn/false_negative": 0.1686291000841043,
"fast_rcnn/fg_cls_accuracy": 0.2782012195121951, "iteration": 5, "loss_box_reg": 0.40475407242774963,
"loss_cls": 1.8559419512748718, "loss_mask": 0.6982034742832184, "loss_rpn_cls": 0.05218731239438057,
"loss_rpn_loc": 0.007734265644103289, "lr": 0.00011990000000000001, "mask_rcnn/accuracy": 0.2920252418802887,
"mask_rcnn/false_negative": 0.9506743610216735, "mask_rcnn/false_positive": 0.2467277275646746,
"roi_head/num_bg_samples": 115.0, "roi_head/num_fg_samples": 13.0, "rpn/num_neg_anchors": 251.25,
"rpn/num_pos_anchors": 4.75, "time": 0.2812129614999961, "total_loss": 2.9480794025585055}

Final: {"data_time": 0.006645713496254757, "eta_seconds": 0.0, "fast_rcnn/cls_accuracy": 0.984375,
"fast_rcnn/false_negative": 0.0, "fast_rcnn/fg_cls_accuracy": 1.0, "iteration": 99999,
"loss_box_reg": 0.054501257836818695, "loss_cls": 0.03908616118133068, "loss_mask": 0.06379510834813118,
"loss_rpn_cls": 0.0010549294238444418, "loss_rpn_loc": 0.005114196799695492, "lr": 0.02,
"mask_rcnn/accuracy": 0.9722937165937804, "mask_rcnn/false_negative": 0.012800597698663846,
"mask_rcnn/false_positive": 0.05609636495338312, "roi_head/num_bg_samples": 115.75,
"roi_head/num_fg_samples": 12.25, "rpn/num_neg_anchors": 246.5, "rpn/num_pos_anchors": 9.5,
"time": 0.3563086399990425, "total_loss": 0.15775199318886735}
\end{verbatim}\todo{Training details, will turn this into a paragraph on how model is created}
\subsection{Pixel Analysis}
The masks and bounding boxes produced by \verb|detectron| are generally quite good, although they are almost never perfect. They almost always either include additional background as part of the fish, or clip part(s) of the fish outside the bounding box. To solve this shortcoming, we utilize pixel analysis methods commonly found in image informatics.

@@

To overcome this, the threshold value needs either scaled up or down depending on if the background is lighter or darker than the fish. For our current dataset, the background is always lighter (i.e.\ closer to 255), so the threshold value needs scaled up to include more of the image as foreground. Precisely how much to scale this value by is difficult to determine. For optimal results it is dependent on the level of contrast between the background and foreground, which can be affected by both how washed out the image is and how light or dark the pigmentation of the fish is. To take this into account, we measure the mean intensity of the foreground and background, then use these intensity values and the difference between them to scale the threshold value. Even with these values in hand it is not entirely clear how to scale the threshold. We scaled the threshold value by 50\% of the difference between the mean of the background and the original threshold value, which captures most of fish's fins in most cases without also masking parts of the background.

@@

\begin{table}[H]
    \centering
    \caption{Metadata properties}
    \label{tab:properties}
    \begin{tabular}{cccp{0.5\linewidth}}
        \toprule
        \textbf{Property} & \textbf{Association} & \textbf{Type} & \textbf{Explanation}\\
        \midrule
        \verb|has_fish| & Overall Image & Boolean & Whether a fish was found in the image.\\
        \verb|fish_count| & Overall Image & Integer & The quantity of fish present.\\
        \verb|has_ruler| & Overall Image & Boolean & Whether a ruler was found in the image.\\
        \verb|ruler_bbox| & Overall Image & 4 Tuple & The bounding box of the ruler (if found).\\
        \verb|scale| & Overall Image & Float & The scale of the image in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|bbox| & Per Fish & 4 Tuple & The top left and bottom right coordinates of what define the bounding box for a fish.\\
        \verb|background.mean| & Per Fish & Float & The mean intensity of the background within a given fish's bounding box.\\
        \verb|background.std| & Per Fish & Float & The standard deviation of the background within a given fish's bounding box.\\
        \verb|foreground.mean| & Per Fish & Float & The mean intensity of the foreground within a given fish's bounding box.\\
        \verb|foreground.std| & Per Fish & Float & The standard deviation of the foreground within a given fish's bounding box.\\
        \verb|centroid| & Per Fish & 4 Tuple & The centroid of a given fish's bitmask.\\
        \verb|primary_axis| & Per Fish & 2D Vector & The unit length primary axis (eigenvector) for the bitmask of a given fish.\\
        \verb|clock_value| & Per Fish & Integer & A fish's primary axis converted into an integer ``clock value'' between 1 and 12.\\
        \verb|length| & Per Fish & Float & The length of a fish in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|mask| & Per Fish & 2D Matrix & The bitmask of a fish in 0's and 1's.\\
        \verb|pixel_analysis_failed| & Per Fish & Boolean & Whether the pixel analysis process failed for a given fish. If \verb|true|, \verb|detectron|'s mask and bounding box were used for metadata generation.\\
        \verb|score| & Per Fish & Float & The percent confidence score output by \verb|detectron| for a given fish.\\
        \verb|has_eye| & Per Fish & Boolean & Whether an eye was found for a given fish.\\
        \verb|eye_center| & Per Fish & 2 Tuple & The centroid of a fish's eye.\\
        \verb|side| & Per Fish & String & The side (i.e.\ \verb|'left'| or \verb|'right'|) of the fish that is facing the camera (dependent on finding its eye).\\
      \bottomrule
\end{tabular}
\end{table}

@@

Pixel analysis failed: 0.036336004421110804

My hand checking of the ''wrong wrong'' fish:
               My guess, Yasin spreadsheet
INHS_FISH_101109.jpg: 9, 11, I'm wrong although it's fairly curved
INHS_FISH_56818.jpg: 4, 2, I'm wrong although there's a tag that is likely messing things up
INHS_FISH_57607.jpg: 3, 9, I'm wrong
INHS_FISH_57001.jpg: 4, 2, I'm wrong although there's a tag that is likely messing things up
INHS_FISH_63953.jpg: 3, 9, I'm wrong
INHS_FISH_51100.jpg: 3, 9, I'm wrong
INHS_FISH_80528.jpg: 3, 8, I'm wrong
INHS_FISH_6370.jpg: 3, 9, I'm wrong
INHS_FISH_54157.jpg: 9, 7, I'm arguably wrong although it's very curved
INHS_FISH_80987.jpg: 9, 3, I'm wrong although this one is a fairly atypical image
